Context Inconsistency Management \\in Pervasive Systems

### 2011-04-07 paper/paper.tex
review for language

In this paper we present two complementary approaches for improving the mitigation of context inconsistencies. 
First we present partial constraint checking for timely identifying context inconsistencies at runtime. An extra constraint layer is added to the traditional ontology based context model and conflicts can be detected by locally checking partial constraints in the ontology. This dramatically improves performance compared to iterative evaluation of an entire ontology~\cite{xu:2010:PCC}.
Secondly we discuss the extention of the traditional ontology model with context lifecycles to more accurately represent the environment of context-aware applications. This information can then be used to estimate the relative reliability of contexts in a conflict set and discard the contexts with lowest reliabilty~\cite{bu:2006:CCM}. 
Apart from resolving context conflicts it is also possible to represent inconsistencies into the context model itself~\cite{ko:2009:IOFO}. In this paper we present a use case in which we explore the possibilities of incorporating inconsitencies into context models using fuzzy ontologies.
} % end of abstract

@@

In pervasive computing, contexts are used to make applications perform a certain action. Contexts can be seen as pieces of environmental information, which the applications depend on. Sometimes contexts are not consistent due to cross reads and misreads. The correctness of the contexts can be checked by evaluating the consistency constraints associated to them. There are multiple ways to check for consistency constraints. Constraint checking techniques are split up in two classes: Non-incremental checking and incremental checking. Incremental checking consists also of two classes: Entire Constraint Checking and Partial Constraint Checking. Once inconsistent contexts have been detected, an algorithm can be used to select the correct contexts. The algorithms use specific data structures, like ontology models, in order to resolve the inconsistencies. However, certain conflict resolving algorithms are too slow for use in practice. This paper describes a way to use an ontology model in combination with the Partial Constraint Checking algorithm in order to detect context inconsistencies fast enough, such that it can be used in practice. Another approach to handling context inconsistencies is representing them into the context model, as opposed to resolving them. This can be done efficiently by using fuzzy ontologies in a tree like structure, as we show in a use case.

\section{Related Work}
Pervasive or ubiquitous computing is a fast-developing discipline that has been receiving increasing attention from both researchers and software developers~\cite{xu:2010:PCC}. In the past decade, many context-aware systems have been developed, ranging from smart room environments to warehouse and supply chain management systems. Lots of effort has been put into building middleware infrastructures that handle vast amounts of sensory data and extract the context information relevant for pervasive applications. Examples of such systems are CoBrA~\cite{1276865} and CORTEX~\cite{1276875}. Various modeling approaches have been proposed for capturing context information, of which the ontology based context model appears to be most promising for most pervasive applications~\cite{bu:2006:CCM}.

Context management for consistency however has not been adequately studied in the existing literature. None of the studies on context-awareness discusses a way for detecting context inconsistencies for reliable pervasive computing~\cite{xu:2010:PCC, bu:2006:CCM}. Even though there has been related research going on in other disciplines as artificial intelligence and software engineering, it doesn't provide adequate support for context inconsistency detection in ubiquitous computing. In addition the strategies proposed in literature for resolving context conflicts are not suited for pervasive computing. Some are based on assumptions that may not apply to general pervasive environments. Other require human participation for conflict resolution, which is usually expensive and slow for pervasive computing~\cite{xu:2010:PCC}. Finally no research has been done on the potential of fuzzy ontologies for representing context inconsistencies in the environment instead of trying to resolve them~\cite{ko:2009:IOFO}.

@@

\section{Partial Constraint Checking}
Constraint checking techniques have been extensively studied in software engineering. Existing constraint checking algorithms focus on checking software artifacts that do not change rapidly or frequently~\cite{xu:2010:PCC}. Context-aware applications require more efficient algorithms because they use a huge set of contexts, which can change very rapidly and frequently (in the range of milliseconds). An inefficient software solution for this does not only require massive computing power, but interestingly also induces a higher inconsistency detection miss rate: because of the computing delay conflicting contexts slip through the context buffer before they are detected by the software~\cite{xu:2010:PCC}. One example of an inefficient approach is non-incremental checking: whenever there is a change in the set of software artifacts these artifacts are each checked against the entire set of consistency constraints to find out all detectable inconsistencies. An improvement to this would be incremental checking: only a subset of all constraints are checked, namely those that are affected by the specific change in the artifact set. But the real merit of Xu et \textit{al.} was replacing the traditional entire constraint checking approach by partial constraint checking based on a consistency computation tree~\cite{xu:2010:PCC}. Their idea is that constraints can be represented as trees with nodes for logical operations and leaves for specific properties of contexts or context sets. Whenever a context is added to or deleted from a context set in time, the branch corresponding to this context can respectively be added to or deleted from the tree (see Fig.~\ref{fig:cct}). 
\begin{figure}[htb]
  \centering
  \includegraphics[width=3.5in]{cons_comp_tree}
  \caption{Consistency Computation Tree: branches can be added or deleted to represent changes in a context set. Source:~\cite{xu:2010:PCC}}
  \label{fig:cct}
\end{figure}
Because intermediate values are retained in the tree nodes after calculation they can be reused whenever the tree changes. More specifically, when a branch is added, only the values for the new branch itself and for the nodes from the branch top to the tree root need to be calculated. When a branch is deleted, only the values for the nodes from the branch top to the tree root have to be recalculated.

With their partial constraint checking algorithm Xu et \textit{al.} attained a time complexity between $O(n)$ for the worst case and $O(1)$ for the best case when a context was added to the set and $O(1)$ when a context was removed. Compared to traditional constraint checking with an overall complexity of $O(n)$, this is a dramatical improvement. In their experiments Xu et \textit{al.} showed that their performance is 15 times better than the traditional approach and in a case study the inconsistency miss rate dropped from $52.2\%$ in traditional checking to $0.1\%$ with partial checking~\cite{xu:2010:PCC}.

@@

\section{Context Inconsistency Resolution}
Now that we have an efficient method for detecting inconsistencies the task of resolving the conflicts remains. As discussed above the strategies for conflict resolution in literature are not very suited for pervasive computing because their assumptions do not apply or they require human participation~\cite{xu:2010:PCC}. What really is needed is an algorithm that in the case of a conflict between two or more contexts decides which context has the highest reliability and retains that context. Bu et \textit{al.} show us that this is possible through extending the ontology based context model with additional information about the context's status and temporal properties~\cite{bu:2006:CCM}. More specifically they propose an algorithm that for every conflict set retains the context with the highest \textit{relative frequency}: frequency of context updates relative to update interval and context age (see Fig.~\ref{fig:cir}). 
\begin{figure}[htb]
  \centering
  \includegraphics[width=3.5in]{cir}
  \caption{Context inconsistency resolution in action. Source:~\cite{bu:2006:CCM}}
  \label{fig:cir}
\end{figure}
This is based on the assumption that contexts that are most stably perceived by a system are most likely to be correct. This assumption is applicable to most of the environments in which pervasive systems run and is applied in many domains as a domain specific approach~\cite{xu:2010:PCC}.

\section{Fuzzy Ontologies}
In pervasive computing traditionally first order logic is used to model the environment of an application. This results in a so called ontology based context model and every observation that does not exactly fit in this model results into a conflict that leads to misadjusted behavior in applications unless they are timely detected and resolved~\cite{xu:2010:PCC}. Because it is impossible to perfectly model the un-predictive environment in which context-aware applications run, it would be nice if the context model was more resilient to unexpected observations. As we illustrate in the case study below, this can be obtained by using fuzzy ontologies.
\subsection{Case Study}
Suppose we want to build a context-aware application that has certain notion of teacher's activities to assist them in the best possible way (for instance automatically opening slides when a lecture starts). In the traditional approach we could build an ontology specifying where lectures take place and that a lecture starts when the teacher appears at the lectern. If now an unpredicted observation is made (sensor cross read or teacher momentarily leaves the room), an inconsistency occurs that has to be resolved for the application to proceed with normal operation. In the fuzzy ontology approach these unpredictable events can simply be modeled so that the ontology is more robust and always yields the most apt interpretation of the environment without conflict resolution. A simplified example of such an ontology is depicted in Fig.~\ref{fig:fuzzy}.
\begin{figure}[htb]
  \centering
  \includegraphics[width=3.5in]{fuzzy}
  \caption{Fuzzy ontology specifying teachers activities and their most probable location.}
  \label{fig:fuzzy}
\end{figure}
Using this fuzzy ontology the relative probability of each specified higher-level context can be calculated from the contexts in the observation window. For instance the observation sequence [$L_1$, $L_2$, $L_3$, $A_1$, $A_2$] gives a relative probability of 0.9998 for `Lecturing' and 0.0002 for `Working', event though the observations are clearly conflicting, since the teacher clearly cannot be at the lectern and in the aisle at the same time. The sequence of [$L_1$, $A_1$, $A_2$, $O_1$, $O_2$] gives a 0.941 probability for `Working' and 0.059 for `Lecturing', even though the teacher strictly spoken cannot be at the lectern and in the office at the same time (see appendix for exact calculations).

@@

Interestingly a fuzzy ontology has some tree-like structure that allows for very efficient incremental partial probability calculation (PPC). The probability of a higher-level context is calculated by simply multiplying the probabilities of the observed contexts. When a new context is added to the observation window, the intermediate calculation result stored at a node (the root in our case study) can simply be multiplied by de probability of the new observation in order to obtain the updated probability for a higher-level context. When a context disappears from the observation window, the intermediate result can simply be divided by the context's probability to obtain an updated probability for a higher-level context.

@@

\section{Discussion}
In this article we started with presenting partial constraint checking: an efficient algorithm for timely detecting inconsistencies in huge sets of contexts which change very frequently. Partial constraint checking was 15 times as efficient as the traditional greedy approach and reduced the conflict detection miss rate from 52.2\% to 0.1\%. This algorithm is clearly a dramatic improvement of the traditional approach and vital to any context aware application that has to handle high-frequently generated contexts.

>>>>

In this paper we present two complementary approaches for improving the mitigation of context inconsistencies. 
First we present partial constraint checking for timely identifying context inconsistencies at runtime. An extra constraint layer is added to the traditional ontology based context model and conflicts can be detected by locally checking partial constraints in the ontology. This dramatically improves performance compared to iterative evaluation of an entire ontology~\cite{xu:2010:PCC}.
Secondly we discuss the extension of the traditional ontology model with context life cycles to more accurately represent the environment of context-aware applications. This information can then be used to estimate the relative reliability of contexts in a conflict set and discard the contexts with lowest reliability~\cite{bu:2006:CCM}. 
Apart from resolving context conflicts it is also possible to represent inconsistencies into the context model itself~\cite{ko:2009:IOFO}. In this paper we present a use case in which we explore the possibilities of incorporating inconsistencies into context models using fuzzy ontologies.
} % end of abstract

@@

In pervasive computing, contexts are used to make applications perform a certain action. Contexts can be seen as pieces of environmental information, which the applications depend on. Sometimes contexts are not consistent due to cross reads and misreads. The correctness of the contexts can be checked by evaluating the consistency constraints associated to them. There are multiple ways to check for consistency constraints. Constraint checking techniques are split up in two classes: Non-incremental checking and incremental checking. Incremental checking also consists of two classes: Entire Constraint Checking and Partial Constraint Checking. Once inconsistent contexts have been detected, an algorithm can be used to select the correct contexts. The algorithms use specific data structures, like ontology models, in order to resolve the inconsistencies. However, certain conflict resolving algorithms are too slow for use in practice. This paper describes a way to use an ontology model in combination with the Partial Constraint Checking algorithm in order to detect context inconsistencies fast enough, such that it can be used in practice. Another approach to handling context inconsistencies is representing them into the context model, as opposed to resolving them. This can be done efficiently by using fuzzy ontologies in a tree like structure, as we show in a use case.

\section{Related Work}
Pervasive or ubiquitous computing is a fast-developing discipline that has been receiving increasing attention from both researchers and software developers~\cite{xu:2010:PCC}. In the past decade, many context-aware systems have been developed, ranging from smart room environments to warehouse and supply chain management systems. A lot of effort has been put into building middleware infrastructures that handle vast amounts of sensory data and extract the context information relevant for pervasive applications. Examples of such systems are CoBrA~\cite{1276865} and CORTEX~\cite{1276875}. Various modeling approaches have been proposed for capturing context information, of which the ontology based context model appears to be most promising for most pervasive applications~\cite{bu:2006:CCM}.

Context management for consistency however has not been adequately studied in the existing literature. None of the studies on context-awareness discusses a way for detecting context inconsistencies for reliable pervasive computing~\cite{xu:2010:PCC, bu:2006:CCM}. Even though other disciplines as artificial intelligence and software engineering conducted relate research, it does not provide adequate support for context inconsistency detection in ubiquitous computing. In addition the strategies proposed in literature for resolving context conflicts are not suited for pervasive computing. Some are based on assumptions that may not apply to general pervasive environments. Other require human participation for conflict resolution, which is usually expensive and slow for pervasive computing~\cite{xu:2010:PCC}. Finally no research has been done on the potential of fuzzy ontologies to represent context inconsistencies in the environment instead of trying to resolve them~\cite{ko:2009:IOFO}.

@@

\section{Partial Constraint Checking}
Constraint checking techniques have been extensively studied in software engineering. Existing constraint checking algorithms focus on checking software artifacts that do not change rapidly or frequently~\cite{xu:2010:PCC}. Context-aware applications require more efficient algorithms because they use a huge set of contexts, which can change very rapidly and frequently (in the range of milliseconds). An inefficient software solution for this does not only require massive computing power, but interestingly also induces a higher inconsistency detection miss rate. Because of the computing delay conflicting contexts slip through the context buffer before they are detected by the software~\cite{xu:2010:PCC}. One example of an inefficient approach is non-incremental checking: whenever there is a change in the set of software artifacts these artifacts are each checked against the entire set of consistency constraints to find out all detectable inconsistencies. An improvement to this would be incremental checking: only a subset of all constraints are checked, namely those that are affected by the specific change in the artifact set. But the real merit of Xu et \textit{al.} was replacing the traditional entire constraint checking approach by partial constraint checking based on a consistency computation tree~\cite{xu:2010:PCC}. Their idea is that constraints can be represented as trees with nodes for logical operations and leaves for specific properties of contexts or context sets. Whenever a context is added to or deleted from a context set in time, the branch corresponding to this context can respectively be added to or deleted from the tree (see Fig.~\ref{fig:cct}). 
\begin{figure}[htb]
  \centering
  \includegraphics[width=3.5in]{cons_comp_tree}
  \caption{Consistency Computation Tree: branches can be added or deleted to represent changes in a context set. Source:~\cite{xu:2010:PCC}}
  \label{fig:cct}
\end{figure}
Because intermediate values are retained in the tree nodes after calculation they can be reused whenever the tree changes. More specifically, when a branch is added, only the values of the new branch itself and of the nodes from the branch top to the tree root need to be calculated. When a branch is deleted, only the values for the nodes from the branch top to the tree root have to be recalculated.

With their partial constraint checking algorithm Xu et \textit{al.} attained a time complexity between $O(n)$ in the worst case and $O(1)$ in the best case when a context is added to the set and $O(1)$ when a context is removed. Compared to traditional constraint checking with an overall complexity of $O(n)$, this is a dramatical improvement. In their experiments Xu et \textit{al.} showed their performance is 15 times better than the traditional approach and in a case study the inconsistency miss rate dropped from $52.2\%$ in traditional checking to $0.1\%$ with partial checking~\cite{xu:2010:PCC}.

@@

\section{Context Inconsistency Resolution}
Now that we have an efficient method for detecting inconsistencies the task of resolving the conflicts remains. As discussed above the strategies for conflict resolution in literature are not well suited for pervasive computing because their assumptions do not apply or they require human participation~\cite{xu:2010:PCC}. What is really needed is an algorithm that in the case of a conflict between two or more contexts decides which context has the highest reliability and retains that context. Bu et \textit{al.} show us that this is possible through extending the ontology based context model with additional information about the context's status and temporal properties~\cite{bu:2006:CCM}. More specifically they propose an algorithm that for every conflict set retains the context with the highest \textit{relative frequency}: frequency of context updates relative to update interval and context age (see Fig.~\ref{fig:cir}). 
\begin{figure}[htb]
  \centering
  \includegraphics[width=3.5in]{cir}
  \caption{Context inconsistency resolution in action. Source:~\cite{bu:2006:CCM}}
  \label{fig:cir}
\end{figure}
This is based on the assumption that contexts that are most stably perceived by a system are most likely to be correct. This assumption is applicable to most of the environments in which pervasive systems run and is applied in many domains as a domain specific approach~\cite{xu:2010:PCC}.

\section{Fuzzy Ontologies}
In pervasive computing traditionally first order logic is used to model the environment of an application. This results in a so called ontology based context model. Every observation that does not exactly fit in this model results into a conflict. This conflict can lead to misadjusted behavior in applications unless they are timely detected and resolved~\cite{xu:2010:PCC}. Because it is impossible to perfectly model the un-predictive environment in which context-aware applications run, it would be nice if the context model was more resilient to unexpected observations. As we illustrate in the case study below, this can be obtained by using fuzzy ontologies.
\subsection{Case Study}
Suppose we want to build a context-aware application that has certain notion of teacher's activities to assist them in the best possible way (for instance automatically opening slides when a lecture starts). In the traditional approach we could build an ontology specifying where lectures take place and that a lecture starts when the teacher appears at the lectern. If now an unpredicted observation is made (sensor cross read or teacher momentarily leaves the room), an inconsistency occurs that has to be resolved for the application to proceed with normal operation. In the fuzzy ontology approach these unpredictable events can simply be modeled so that the ontology is more robust and always yields the most apt interpretation of the environment without conflict resolution. A simplified example of such an ontology is depicted in Fig.~\ref{fig:fuzzy}.
\begin{figure}[htb]
  \centering
  \includegraphics[width=3.5in]{fuzzy}
  \caption{Fuzzy ontology specifying teachers activities and their most probable location.}
  \label{fig:fuzzy}
\end{figure}
Using this fuzzy ontology the relative probability of each specified higher-level context can be calculated from the contexts in the observation window. For instance the observation sequence [$L_1$, $L_2$, $L_3$, $A_1$, $A_2$] gives a relative probability of 0.9998 for `Lecturing' and 0.0002 for `Working', event though the observations are clearly conflicting, since the teacher clearly cannot be at the lectern and in the aisle at the same time. The sequence of [$L_1$, $A_1$, $A_2$, $O_1$, $O_2$] gives a 0.941 probability for `Working' and 0.059 for `Lecturing', even though the teacher strictly spoken cannot be at the lectern and in the office at the same time (see appendix for exact calculations).

@@

Interestingly a fuzzy ontology has some tree-like structure that allows for very efficient incremental partial probability calculation (PPC). The probability of a higher-level context is calculated by simply multiplying the probabilities of the observed contexts. When a new context is added to the observation window, the intermediate calculation result stored at a node (the root in our case study) can simply be multiplied by the probability of the new observation in order to obtain the updated probability for a higher-level context. When a context disappears from the observation window, the intermediate result can simply be divided by the context's probability to obtain an updated probability for a higher-level context.

@@

\section{Discussion}
In this article we started with presenting partial constraint checking: an efficient algorithm for timely detecting inconsistencies in huge sets of contexts which change frequently. Partial constraint checking was 15 times as efficient as the traditional greedy approach and reduced the conflict detection miss rate from 52.2\% to 0.1\%. This algorithm is clearly a dramatic improvement of the traditional approach and vital to any context aware application that has to handle frequently generated contexts.
